Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-04 paper.tex
Comments on reproducibility section

             Containeri
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        Reproducibility is a critical component to all scientific workflows. 
        Without reproducibility, correctness of results could never be 
        determined. Reproducibility is defined as the ability to obtain similar 
        results by 
        recreating or executing the original workflow on the same or different
        infrastructure\note{cite Peng 2011}. It has been found that more than 70\% of
        scientific experiments are not reproducible \note{cite Mbaker}

        Current studies show that different tools 
        performing the same function, in addition to the same versions of tools 
        executing on top of different operating systems, produce significantly 
        dissimilar results \cite{bowring, gronenschild 2011, glatard 2015}. As a result, it is necessary that specific details 
        on the workflow, such a data used, software versions and operating 
        system used are provided. For this reason, many workflow engines have
        incorporated provenance tracking into their design.

@@

        This chapter will examine how different workflow engines tackle the 
        issue of reproducibility. Section~\ref{metadata} will investigate three 
        different provenance metadata specifications available (W3C PROV, NIDM,
        and BIDS Derivates).Section~\ref{prov} will discuss how the 
        engines assimilate and distribute the provenance data. 
        Section~\ref{repcont} will examine how these engines use containers to
        facilitate reproducibility of the experiments.
        \section{Metadata formats}\label{metadata}
            The recent increase in data-sharing resulted in the rise of BigData 
            in neuroscience. To continue the promotion of data sharing in 
            neuroscience, the derived data's metadata must be well understood
            by the community. Even if the provenance tracking is comprehensive,
            it is useless if it is only understood by some. To address this 
            issue, specifications for the standardization of metadata have been 
            proposed. Here we look at three provenance metadata specifications.
            The first is W3 PROV, a general-purpose provenance specification; 
            NIDM-results
            a comprehensive neuroimaging provenance specification; and BIDS 
            Derivatives the recent provenance addition to the family of 
            popular neuroimaging metadata specifications, BIDS.

            \subsection{W3C PROV}
            The W3C PROV specification came out from the need to explain the 
            origin of data shared on the internet. 
            The aim of the W3 PROV specification was gather information on
            the activities, entities and involved parties in the creation of 
            the data. This metadata would allow for assements of the quality, 
            reliability and trustworthiness to be performed. Sharing of the 
            provenance is faciliated through the use of widely-used formats 
            such as RDF and XML.

            The PROV model has three main objects involved: 1) entities, 2) 
            activities and 3) agents. Any kind of thing is considered an entity
            . Actions act upon entities to generate new ones. Activities have 
            agents associated to them. These agents are responsible for the 
            activity taking place. Agents can be individuals, companies and 
            software, amongst others.

            \subsection{NIDM-Results}
                NIDM-results was designed to improve the limited XML-based 
                Clinical and Experimental Data Exchange (XCEDE) neuroimaging
                provenance metadata format. It is a domain-specific extension
                to the W3 PROV model that aims to provide comprehensive and
                machine-readable provenance from raw data to its final stage.
                However, it is currently limited in the provenance 
                information it can provide.

            \subsection{BIDS Derivatives}
                The BIDS-Derivatives metadata specification is the newest 
                addition to the BIDS family of neuroimaging-related 
                specifications. It is the sister
                specification to the widely adopted BIDSRaw metadata format 
                and it remains compatible with the BIDS Raw specification. 
                Unlike NIDM-Results, BIDS derivatives, like BIDS Raw, do not
                use the W3C PROV specification. Rather, the provenance can be
                derived from its folder structure and filename. Any additional
                metadata can be found in a JSON sidecar file.

@@

            As explained by the W3 PROV specification, provenance details 
            information on the data, the tools involved in its creation, and 
            the responsible parties. Provenance tracking, in the context of 
            workflows, is the act of capturing provenance data during
            workflow execution. With detailed provenance and access to source 
            code information, researchers should have enough information to be 
            able to reproduce workflows.

@@

            FastR performs provenance very similarly to Nipype. However,
            FastR's agents can be Fastr, Networks, Tools, Nodes. Additionally,
            activities, in
            FastR map to Jobs and entities map to the data.

@@

            While all workflow engines capture generally the same type of 
            provenance information, they differ in how much provenance 
            information is captured and how it's done. LONI, Pegasus, Pydpiper 
            and Galaxy
            capture an above minimum amount of provenance data, whereas engines 
            like Nextflow, Make, Toil and MapReduce capture very limited  
            provenance 
            information. OpenMOLE handles provenance through encapsulation of
            tasks using CARE to simulate runtime environments
            and BigData frameworks like Spark and Dask mainly rely on 
            lineage.

            An above minimum amount of provenance details captured includes 
            capturing details on the workflow stages as well as the inputs and
            outputs. LONI stores this type of information in an XML file.

@@

            To handle issues with numerical instability of tools and relieve 
            users from
            having to configure their environment to be almost identical to 
            that of the study they are trying to replicate, significant 
            emphasis has been placed on containerization. Similarly to 
            OpenMOLE's CARETasks, containerization provides an encapsulated
            environment for the tools to run which mimics the environment of 
            the original study. This ensures that users executing their 
            workflows in the same environments will obtain the same results.
            A major issue with containerization is that despite it controlling
            environments and permitting experiments to be reproducible, it does
            not solve the issue of numerical instability. Moreover, this 
            mentality may lead researchers to limit themselves to older tool 
            versions and further limit scientific advancement.

>>>>

             Containeri
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        Reproducibility is a critical component to all scientific workflows. 
        Without reproducibility, correctness of results could never be 
        determined \tristan{Link between two first sentences could be smoothened}. Reproducibility is defined as the ability to obtain similar 
        results by 
        recreating or executing the original workflow on the same or different
        infrastructure\note{cite Peng 2011}. It has been found that more than 70\% of
        scientific experiments are not reproducible \note{cite Mbaker}

        Current studies show that different tools 
        performing the same function, in addition to the same versions of tools 
        executing on top of different operating systems, produce significantly 
        dissimilar results \cite{bowring, gronenschild 2011, glatard 2015}. As a result, it is necessary that specific details 
        on the workflow, such a data used, software versions and operating 
        system used are provided. For this reason, many workflow engines have
        incorporated provenance tracking into their design. \tristan{Provenance is also
        a way to merely document what happened so that results can be traced.}

@@

        This chapter will examine how different workflow engines tackle the 
        issue of reproducibility. Section~\ref{metadata} will investigate three 
        different provenance specifications used in neuroinformatics (W3C PROV, NIDM,
        and BIDS Derivates) \tristan{I don't think you want to talk about 'metadata'. Just
        mention 'provenance'.}. Section~\ref{prov} will discuss how the 
        engines assimilate and distribute the provenance data. 
        Section~\ref{repcont} will examine how these engines use containers to
        facilitate reproducibility of the experiments.
        \section{Provenance Formats}\label{metadata}
            The recent increase in data-sharing resulted in the rise of BigData 
            in neuroscience \tristan{This feels like general intro to me}. To continue the promotion of data sharing in 
            neuroscience, the derived data's metadata must be well understood
            by the community. Even if the provenance tracking is comprehensive,
            it is useless if it is only understood by some. To address this 
            issue, specifications for the standardization of metadata have been 
            proposed. Here we look at three provenance metadata specifications.
            The first is W3 PROV, a general-purpose provenance specification; 
            NIDM-results
            a comprehensive neuroimaging provenance specification; and BIDS 
            Derivatives the recent provenance addition to the family of 
            popular neuroimaging metadata specifications, BIDS \tristan{Make sure that
            the acronym is defined earlier.}.

            \subsection{W3C PROV}
            The W3C PROV specification came out from the need to explain the 
            origin of data shared on the Web. 
            The aim of the W3 PROV specification was \tristan{to?} gather information on
            the activities, entities and involved parties in the creation of 
            data. This metadata would allow for assessments of the quality, 
            reliability and trustworthiness to be performed. Sharing of the 
            provenance is faciliated through the use of widely-used formats 
            such as RDF and XML.

            The PROV model has three main objects involved: 1) entities, 2) 
            activities and 3) agents. Any kind of thing is considered an entity.
            Actions act upon entities to generate new ones. Activities have 
            agents associated to them. These agents are responsible for the 
            activity taking place. Agents can be individuals, companies and 
            software, amongst others.

            \subsection{NIDM-Results}
                NIDM-results was designed to improve the limited XML-based 
                Clinical and Experimental Data Exchange (XCEDE) neuroimaging
                provenance metadata format \tristan{I guess you'll cite references for that}. It is a domain-specific extension
                to the W3 PROV model that aims to provide comprehensive and
                machine-readable provenance from raw data to its final stage. \tristan{give example?}
                However, it is currently limited in the provenance 
                information it can provide. \tristan{Be more specific, what are the limitations?}

            \subsection{BIDS Derivatives}
                The BIDS-Derivatives metadata specification is the newest 
                addition to the BIDS family of neuroimaging-related 
                specifications. It is the sister
                specification to the widely adopted BIDSRaw metadata format 
                and it remains compatible with the BIDS Raw specification. 
                Unlike NIDM-Results, BIDS derivatives, like BIDS Raw, do not
                use the W3C PROV specification. Rather, the provenance can be
                derived from its folder structure and filename. Any additional
                metadata can be found in a JSON sidecar file. \tristan{Explain that
                although BIDS doesn't rely on semantic web technologies, this is the
                solution being adopted in the community, mostly for simplicity. Mention
                that there is a community effort around it and that bidsraw has been
                very successful.}

            \tristan{You might want to add data lineage here.}

@@

            As explained by the W3 PROV specification, provenance details 
            information on the data, the tools involved in its creation, and 
            the responsible parties. Provenance tracking, in the context of 
            workflows, is the act of capturing provenance data during
            workflow execution. With detailed provenance and access to source 
            code information, researchers should have enough information to be 
            able to reproduce workflows. \tristan{These 2 paragraphs should be in the chapter intro.}

@@

            FastR performs provenance very similarly to Nipype. However,
            FastR's agents can be Fastr, Networks, Tools, Nodes. Additionally,
            activities, in
            FastR map to Jobs and entities map to the data \tristan{how is the mapping in nipype? If different,
            it might be a good place to comment on the fact that PROV can be used
            in many different ways}.

@@

            While all workflow engines capture generally the same type of 
            provenance information, they differ in how much provenance 
            information is captured and how it's done. LONI, Pegasus, Pydpiper 
            and Galaxy
            capture an above minimum amount of provenance data, whereas engines 
            like Nextflow, Make, Toil and MapReduce capture very limited  
            provenance 
            information. OpenMOLE handles provenance through encapsulation of
            tasks using CARE to simulate runtime environments
            and BigData frameworks like Spark and Dask mainly rely on 
            lineage \tristan{which is?}.

            \tristan{The rest of this section is very difficult to read because
            it lacks structure. Perhaps you could classify provenance tracking in PROV-based, lineage-based,
            and other custom formats?}
            An above minimum amount of provenance details captured includes 
            capturing details on the workflow stages as well as the inputs and
            outputs. LONI stores this type of information in an XML file.

@@

            To handle issues with numerical instability of tools and relieve 
            users from
            having to configure their environment to be almost identical to 
            that of the study they are trying to replicate, significant 
            emphasis has been placed on containerization. Similarly to 
            OpenMOLE's CARETasks, containerization provides an encapsulated
            environment for the tools to run which mimics the environment of 
            the original study. This ensures that users executing their 
            workflows in the same environments will obtain the same results.
            A major issue with containerization is that despite it controlling
            environments and permitting experiments to be reproducible, it does
            not solve the issue of numerical instability. Moreover, this 
            mentality may lead researchers to limit themselves to older tool 
            versions and further limit scientific advancement.
            \tristan{Nextflow should be mentioned here, as they make a big deal
            of using containers for reproducibility. I would structure this section in 
            2 paragraphs: (1) what engines support containers, (2) why containers help
            reproducibility and why there aren't enough.}
